---
title: "Homework 1"
author: Rachel Heise
output: github_document
---

Homework 1 Assignment

# Problem 1

Start by calling the tidyverse package.
```{r}
library(tidyverse)
```
Create a data frame that will be used to calculate mean values.
```{r}
mean_df = tibble(
  num_vector = rnorm(n = 10),
  logical_vector = c(num_vector > 0),
  char_vector = c('this', 'is', 'a', 'character', 'vector', 'of', 'length', 'ten', 'called', 'char_vector'),
  factor_vector = factor(c('cat', 'dog', 'fish', 'cat', 'cat', 'fish', 'dog', 'dog', 'dog', 'fish'))
)
```
Check whether a mean can be calculated for each column of the data frame.
```{r}
mean(mean_df$num_vector)
mean(mean_df$logical_vector)
mean(mean_df$char_vector)
mean(mean_df$factor_vector)
```

A mean can be calculated for the standard normal column and for the logical vector, but not for the character vector or the factor vector.


Here, the logical, character, and factor vectors are converted to numeric vectors.
```{r}
as.numeric(mean_df$logical_vector)
```
The logical vector returns 0 and 1 values.
```{r}
as.numeric(mean_df$char_vector)
```
The character vector returns NA values (not available).
```{r}
as.numeric(mean_df$factor_vector)
```
The factor vector returns values of 1, 2, and 3, based on which factor was selected. However, you are not able to take a mean of a factor vector because a value returned for this would have no meaning.

```{r}
log_to_num = as.numeric(mean_df$logical_vector)
log_to_num * mean_df$num_vector

log_to_fact = as.factor(mean_df$logical_vector)
log_to_fact * mean_df$num_vector

log_fact_num = as.numeric(log_to_fact)
log_fact_num * mean_df$num_vector
```

# Problem 2




```{r}

data("penguins", package = "palmerpenguins")

nrow(penguins)
ncol(penguins)

mean_flip_length = mean(penguins$flipper_length_mm, na.rm = TRUE)
```

The Penguins data set contains `r ncol(penguins)` variables. Some of the most useful ones for consideration are `r names(penguins)[5]` which has a mean value of `r round(mean(penguins$flipper_length_mm, na.rm = TRUE), digits = 2)` and a standard deviation of `r round(sd(penguins$flipper_length_mm, na.rm = TRUE), digits = 2)`.

```{r}
library(ggplot2)

penguin_scatterplot = ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + geom_point()

ggsave("penguin_scatterplot", device = "png")

```