---
title: "Homework 1"
author: Rachel Heise
output: github_document
---

Homework 1 Assignment

## Problem 1

Start by loading all the required libraries.

```{r libraries}
library(tidyverse)
library(ggplot2)
```

Create a data frame that will be used to calculate mean values.

```{r mean_dataframe}
mean_df <-
  tibble(
    num_vector = rnorm(10),
    logical_vector = num_vector > 0,
    char_vector = c("a", "b", "c", "d", "e", "f", "g", "h", "i", "j"),
    factor_vector = factor(c("cat", "dog", "fish", "cat", "cat", "fish", "dog", "dog", "dog", "fish"))
  )
```

Check whether a mean can be calculated for each column of the data frame.

```{r means}
mean(pull(mean_df, num_vector))
mean(pull(mean_df, logical_vector))
mean(pull(mean_df, char_vector))
mean(pull(mean_df, factor_vector))
```

A mean can be calculated for the numerical vector and for the logical vector, but not for the character vector or the factor vector. Type coercion occurs for the logical vector which is important to understand.


Next, the logical, character, and factor vectors are converted to numeric vectors.

```{r as_numeric}
as.numeric(pull(mean_df, logical_vector))
as.numeric(pull(mean_df, char_vector))
as.numeric(pull(mean_df, factor_vector))
```

The logical vector returns values of 0 and 1.
The character vector returns NA values (not available).
The factor vector returns values of 1, 2, and 3.


However, you are not able to take a mean of a factor vector because a value returned for this would have no meaning.

```{r logical_transformation}
as.numeric(pull(mean_df, logical_vector)) * pull(mean_df, num_vector)

as.factor(pull(mean_df, logical_vector)) * pull(mean_df, num_vector)

as.numeric(as.factor(pull(mean_df, logical_vector))) * pull(mean_df, num_vector)
```


## Problem 2


```{r data_exploration}

data("penguins", package = "palmerpenguins")

nrow(penguins)
ncol(penguins)

mean_flip_length = mean(penguins$flipper_length_mm, na.rm = TRUE)
```

The Penguins data set contains `r ncol(penguins)` variables. Some of the most useful ones for consideration are `r names(penguins)[5]` which has a mean value of `r round(mean(penguins$flipper_length_mm, na.rm = TRUE), digits = 2)` and a standard deviation of `r round(sd(penguins$flipper_length_mm, na.rm = TRUE), digits = 2)`.

```{r create_scatterplot}

ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + geom_point()

ggsave("penguin_scatterplot", device = "png")

```