---
title: "Homework 1"
author: Rachel Heise
output: github_document
---

P8105 Homework 1

## Problem 1

Start by loading all the required packages.

```{r libraries}
library(tidyverse)
library(ggplot2)
```

Create a data frame that will be used to calculate mean values.

```{r mean_dataframe}
mean_df <-
  tibble(
    num_vector = rnorm(10),
    logical_vector = num_vector > 0,
    char_vector = c("a", "b", "c", "d", "e", "f", "g", "h", "i", "j"),
    factor_vector = factor(c("cat", "dog", "fish", "cat", "cat", "fish", "dog", "dog", "dog", "fish"))
  )
```

Check whether a mean can be calculated for each vector in the data frame.

```{r means}
mean(pull(mean_df, num_vector))
mean(pull(mean_df, logical_vector))
mean(pull(mean_df, char_vector))
mean(pull(mean_df, factor_vector))
```

A mean can be calculated for the numerical vector and for the logical vector, but not for the character vector or the factor vector. Type coercion occurs for the logical vector which is important to understand.


Next, the logical, character, and factor vectors are converted to numeric vectors.

```{r as_numeric, results="hide", warning=FALSE}
as.numeric(pull(mean_df, logical_vector))
as.numeric(pull(mean_df, char_vector))
as.numeric(pull(mean_df, factor_vector))
```

The logical vector returns values of 0 and 1, which is the numeric version of this Boolean, which explains why you can take the mean of a logical vector and get a value between 0 and 1.
The character vector returns NA (not available) values because letters or words cannot be converted to numbers.
Finally, the factor vector returns values of 1, 2, and 3, because R orders the factors numerically. However, you are not able to take a mean of a factor vector because a value returned for this would have no meaning or practical use.

Convert the logical vector to various types and multiply by the numeric vector.

```{r logical_transformation}
as.numeric(pull(mean_df, logical_vector)) * pull(mean_df, num_vector)

as.factor(pull(mean_df, logical_vector)) * pull(mean_df, num_vector)

as.numeric(as.factor(pull(mean_df, logical_vector))) * pull(mean_df, num_vector)
```

The conversion from a logical vector to a factor results in TRUE/FALSE values which cannot be multiplied by the numeric vector.

However, converting the logical vector to a factor vector and then to a numeric vector does allow you to multiply by a numeric vector.

## Problem 2


```{r data_exploration}

data("penguins", package = "palmerpenguins")

```

The Penguins data set contains `r ncol(penguins)` variables and `r nrow(penguins)` observations. Several are quite useful for consideration: `r names(penguins)[5]` has a mean value of `r round(mean(pull(penguins, flipper_length_mm), na.rm = TRUE), digits = 2)` and a standard deviation of `r round(sd(pull(penguins, flipper_length_mm), na.rm = TRUE), digits = 2)`. `r names(penguins)[3]` has a mean value of `r round(mean(pull(penguins, bill_length_mm), na.rm = TRUE), digits = 2)` and a standard deviation of `r round(sd(pull(penguins, bill_length_mm), na.rm = TRUE), digits = 2)`. There are `r nlevels(pull(penguins, species))` species, including `r levels(pull(penguins, species))`. Finally, the years for these penguins range from `r min(pull(penguins, year))` to `r max(pull(penguins, year))`.

```{r create_scatterplot, warning=FALSE}

print(ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + geom_point())

ggsave("penguin_scatterplot.png")

```